### 2018-09-30 paper/finitecomplete-failures.tex
concern about a proof

\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points, described in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa})and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event.

{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$, any
adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$ with
$s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$, and the $fail$ event
cannot occur.

\medskip
\noindent {\bf Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
that $U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, so
$\minhits(n) = \minhits(P/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$
leads always to a PASS verdict and is taken if $\minhits(n) = \varnothing$.
Branch (\ref{eq:ufc}) always leads to test continuation without a verdict.
For the last branch, we note that it can only be entered if $\minhits(P/s) =
\minhits(n) \neq\varnothing$. In this case, any selected minimal hitting set
$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
events from $H\in\minhits(n)$, and the test terminates with PASS event
$\epass$. Note that this argument requires that $Q$ is free of livelocks,
because otherwise the $\epass$-events might not become visible, due to
unbounded sequences of hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
respectively:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of the two cases that at least one test
execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
   $\efail$-event or without giving any verdict. The latter case is also
interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.

\medskip
\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
a trace of every process. In this case, $s$ is also a trace of the product
graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and $G/s =
(G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the
construction of $G$,  we know that $G$ has at most $pq$ reachable states,
because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. By
Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
function of $G$ implies that $u$ is also a trace of $P$ and $Q$, which means
that $(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test $U_F(pq-1)$ accepts all
traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by
construction and by Lemma~\ref{lemma:ufproperties}, $U_F(pq-1)/u =
U_F(pq-1,\#u,G(P)/u)$. Since $s.e\not\in\trc(P)$, $e$ is an element of
$\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. Hence, in at least one execution,
$U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with this
event $e$, so that the test fails. Again, the assumption of non-divergence of
$Q$ is needed for this conclusion.

>>>>

\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa})and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event. This case would also interpreted as FAIL, since then
$\epass\then\Stop$ is not failures-refined by $(Q\parallel[\Sigma]
U_F(k))\hide \Sigma$.
% alcc: part of the explanation above was in the following lemma, but it is already
% relevant here.

\paragraph{Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
\trc(P)$, any adaptive test $U_F(p)$ running in parallel with $Q$ will always
enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
external choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$
with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$, and the $fail$ event
cannot occur.

\paragraph{Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
that $U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, so
$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
$U_F(p,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
\varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
without a verdict. \fixme{I think we need a stronger argument here, as to why
it cannot deadlock.} For the last branch, we note that it can only be entered
if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
minimal hitting set $H\in\minhits(n)$ has a non-empty intersection with each
of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ never blocks
when offered events from $H\in\minhits(n)$, and the test terminates with
event $\epass$. Note that this argument requires that $Q$ is free of
livelocks, because otherwise the $\epass$-events might not become visible,
due to unbounded sequences of hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the $\efail$
event or deadlocks.

\medskip
\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
a trace of every process. In this case, $s$ is also a trace of the product
graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and $G/s =
(G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the
construction of $G$,  we know that $G$ has at most $pq$ reachable states,
because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. By
Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
function of $G$ implies that $u$ is also a trace of $P$ and $Q$, which means
that $(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test $U_F(pq-1)$ accepts all
traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by
construction and by Lemma~\ref{lemma:ufproperties}, $U_F(pq-1)/u =
U_F(pq-1,\#u,G(P)/u)$. Since $s.e\not\in\trc(P)$, $e$ is an element of
$\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. Hence, in at least one execution,
$U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with this
event $e$, so that the test fails. Again, the assumption of non-divergence of
$Q$ is needed for this conclusion. \fixme{There is an assumption here that
the graph of a test execution is given by the cross-product of graphs.  But,
test execution is not defined in this way. It is defined by parallelism and
hiding. I think the definition of test execution needs to change.}
